### 2013-10-02 chi2014/qualitative.tex
table

The second category of strategies is \emph{riffing}. We found riffing manifested itself many ways, and identified six distinct riffing strategies used by participants: \emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:

@@

Finally, \emph{Follow-up riffing} occurs when a participants response assumes that the solution in one of their previous responses has already been implemented. For example:

@@

Follow-up riffing, while interesting, is incredibly rare in the data and we will not discuss it further.

There are two special variants of riffing that can occur across all riffing strategies. Some participants \emph{pair} their riffs, such that they often riff on old ideas but never produce more than two versions of each idea.

Another special case is is \emph{linguistic riffing}. Some participants seem to use common sentence structure as queues to help them generate ideas. For example:

\begin{itemize}
    \item Make sunglasses out of them
    \item Make shoes out of them
    \item Make a handbag out of them
\end{itemize}

Note that linguistic riffing and exhaustive mutation have significant overlap. A participant that is employing exhaustive mutation is always employing linguistic riffing, but not vice-versa.

Finally, \emph{reaching back} is a variant of riffing employed by almost all participants who provided more than 10 responses. Reaching back entails riffing on an idea that occured further back in the run than the previous response. The presence of reaching back speaks to the exaustion of the participant's creativity, as they produce new variants of old ideas rather than new ideas altogether.

Almost all participants employ some form of riffing. This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas". However, the different sub-strategies of riffing have very different impacts on the qualitative assessment of a run. Exhaustive mutation is often unsatisfying, as it feels as though the participant is trying to wring every scrap of task completion out of a single idea. On the other hand, exploratory riffing can produce answers that are incredibly unexpected but still of high quality. Riffing strategies paint a significant picture of the persona of the participant.

@@

Six personas were identified: quitters, lazy creatives, diligent struggles, lazy strugglers, hard workers, and over-deliverers.

@@

>>>>

The second category of strategies is \emph{riffing}.
We found riffing manifested itself many ways, and identified four distinct riffing strategies used by participants: specificity riffing, exhaustive mutation, exploratory riffing, and follow-up riffing. We describe each of these in detail, and then touch on three common phenomena found under riffing strategies.

\emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:

@@

Finally, \emph{Follow-up riffing} occurs when a participant's response assumes that the solution in one of their previous responses has already been implemented. For example:

@@

Within and across these strategies, there are a few repeated phenomena.
Some participants \emph{pair} their riffs. These participants  often riff on old ideas but never produce more than two versions of each idea. This may represent a desire to generate more ideas to meet the task expectations, but a trepitation that repeated ideas are less satisfying than completely unique ones.

Another phenomenon is \emph{linguistic riffing}. Some participants seem to use common sentence structure as queues when generating ideas. For example:

\begin{itemize}
    \item We could use the hard-drives inside for different electronics.
    \item We could use them in place of rocks (to throw at things, to use in pavement.)
    \item We could use them to back up our music collections on our current devices.
\end{itemize}

Note the distinction from exhaustive mutation. In the exhaustive mutation example above, instances using the same language were interchangeable.
In this example, while language is re-used, each instance is a new idea.

Finally, \emph{reaching back} is a variant of riffing employed by almost all participants who provided more than 10 responses. Reaching back entails riffing on an idea that occured further back in the run than the previous response. The presence of reaching back speaks to the exhaustion of the participant's creativity, as they produce new variants of old ideas rather than new ideas altogether.

Almost all participants employ some form of riffing. This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas". However, the different strategies of riffing have very different impacts on the qualitative assessment of a run. Exhaustive mutation is often unsatisfying, as it feels as though the participant is trying to wring every scrap of task completion out of a single idea. On the other hand, exploratory riffing can produce answers that are unexpected and of high quality.

\subsubsection{Time-saving}

@@

Six personas were identified: quitters, lazy creatives, diligent struggles, lazy strugglers, hard workers, and over-deliverers. These personas are described below. A summary, in terms of strategies used, is in Table TAB.

@@

\begin{table*}
    \begin{tabular}{|l|l|l|}
        \hline
        \textbf{Persona} & \textbf{Common strategies} & \textbf{Uncommon strategies} \\
        \hline
        Quitters & cop-outs, exhaustive mutation & \\
        Lazy creatives & & \\
        Diligent strugglers & & \\
        Lazy struggles & & \\
        Hard workers & & \\
        Over-deliverers & & \\
        \hline
    \end{tabular}
\end{table*}
